

---
title: ' Reconciling empirical interactions\ and species coexistence'
output:
  md_document:
    variant: markdown_github  
  pdf_document: default
  html_document: default
---

##

#### Daniel S. Maynard, J. Timothy Wootton, Carlos A. Serván, Stefano Allesina

##

This markdown file provides a brief walk-through of the different functions and methods presented in the manuscript. In this repo there are two main files, ```fitting_algae.R``` and ```fitting_intertidal.R``` that provide step-by-step code for reproducing the results in the main text, along with a ancillary functions called within these scripts (```functions_algae.R``` and ```functions_intertidal.R```). 

##

### Zero-sum replicator example

First, source the R files containing the fitting functions:

```{r echo=T, message=F}
rm(list=ls())
library(limSolve)
library(Matrix)
source("../code/functions_intertidal.R")
source("../code/functions_algae.R")
```

Then, set the seed and the number of species

```{r echo=T, message=F}
set.seed(10)
nspp <- 10
```

Generate a skew-symmetric payoff matrix
```{r echo=T, message=F}
P <- matrix(runif(nspp^2),nspp,nspp)
P <- P-t(P)
round(P,2)
```

And generate a relative abundance vector:
```{r echo=T, message=F}
x_equil <- rexp(nspp)
x_equil <- x_equil/sum(x_equil)
round(x_equil,2)
```


First, let's implement quadratic programming to reconcile P with x_equil, assuming each entry is weighted equally:
```{r echo=T, message=F}
result_zero_sum <- find_closest_matrix_weighted(P=P, x=x_equil)
Pfit <- convert_sol_to_matrix(result_zero_sum$X,nspp)
```

We can view the best-fitting payoff matrix
```{r echo=T, message=F}
round(Pfit,2)
```

And compare to the original matrix
```{r echo=T, message=F}
plot(Pfit~P)
```

Last, we can double check that it is indeed an exact an exact solution by checking the growth rates at x_equil:
```{r echo=T, message=F}
round(Pfit%*%x_equil,12)
```

#### Weighted quadratic programming 

Now let's repeat the above analysis, but this time generate a random matrix of weights, reflecting, for example, different sample sizes for each entry:
```{r echo=T, message=F}
weight_mat <- matrix(runif(nspp^2),nspp,nspp)
round(weight_mat,2)
```
We can provide this weight matrix as an optional argument to *find_closest_matrix_weighted*:
```{r echo=T, message=F}
result_zero_sum_w <- find_closest_matrix_weighted(P=P, x=x_equil, weight_mat = weight_mat)
Pfit_w <- convert_sol_to_matrix(result_zero_sum_w$X,nspp)
```

And we can compare this new weighted payoff matrix to the original best-fitting matrix:

```{r echo=T, message=F}
plot(Pfit~Pfit_w)
```

And to the empirical matrix:
```{r echo=T, message=F}
plot(P~Pfit_w)
```

And once again, it's an exact solution:
```{r echo=T, message=F}
round(Pfit_w%*%x_equil, 12)
```

### Lotka Volterra Example

To fit Lotka-Volterra, we need an estimate of the growth rates and the interaction matrix. So first generate growth rates (r), interaction matrix (A), and a species abundance vector (x_obs):

```{r echo=T, message=F}
r <- runif(nspp)
A <- -matrix(runif(nspp^2), nspp,nspp)
diag(A) <- diag(A)*2
x_obs <- runif(nspp)
round(A,2)
round(r,2)
round(x_obs,2)
```

Instead of using "find_closest_matrix_weighted", as we did for the payoff matrix above, here we using the function "fit_qp_LV". We assign a large tolerance (10,000% deviation possible) allowing each entry to vary as needed to obtain the optimal solution:
```{r echo=T, message=F}
result_LV <- fit_qp_LV(A=A,r=r,x_obs=x_obs,tol=1000)
```

And we convert the solution into the fitted matrix and growth rate vector:
```{r echo=T, message=F}
Afit <- t(matrix(result_LV$X[1:nspp^2], nspp,nspp))
rfit <- result_LV$X[(nspp^2+1):(nspp^2+nspp)]
round(Afit,2)
round(rfit,2)
```

We can compare these fitted parameters to their empirical counterparts:
```{r echo=T, message=F}
plot(Afit~A)
plot(rfit~r)
```

Lastly, we can double check that the fitted values are indeed exact solutions:
```{r echo=T, message=F}
round(x_obs*(rfit+Afit%*%x_obs),12)
```

#### Constraining the entries to be within a fixed percentage of the observed values

The *fit_qp_LV* function can be used to search for a solution within a fixed radius around each empirical value

Here, let's set the tolerance to within +/- 100% of the observed values by specifying tol=1.00
```{r echo=T, message=F}
result_LV_100 <- fit_qp_LV(A=A,r=r,x_obs=x_obs,tol=1.00)
```	

And we can extract the fitted interaction matrix and growth rate vector:
```{r echo=T, message=F}
Afit_100 <- t(matrix(result_LV_100$X[1:nspp^2], nspp,nspp))
rfit_100 <- result_LV_100$X[(nspp^2+1):(nspp^2+nspp)]
round(Afit_100,2)
round(rfit_100,2)
```

Now compare these fitted objects to the originals. Note that it sets a bunch of interactions to be zero:
```{r echo=T, message=F}
plot(Afit_100~A)
plot(rfit_100~r)
```

Lastly, we can compare the two fits, both with and without the constraints:
```{r echo=T, message=F}
plot(Afit_100~Afit)
```

And once again verify that the constrained parameters provide an exact solution:
```{r echo=T, message=F}
round(x_obs*(rfit_100+Afit_100%*%x_obs),12)
```

Under the hood, the functions *fit_qp_LV* and *find_closest_matrix_weighted* both call the function *lsei* from the ```limSolve``` package to implement the quadratic programming step. Please see the *lsei* documentation for more details, and also take a look at the functions *fit_qp_LV* and *find_closest_matrix_weighted* in the files ```functions_algae.R``` and ```functions_intertidal.R```, respectively, to see exactly how these constraints are formulated and implemented. 
